\label[frameworks]
\chap Deep Learning Frameworks

There are tens of deep learning frameworks.
This chapter makes overview of such frameworks in order to choose the
most suitable for purposes of this thesis.
Several aspects have been taken into account:

\begitems
* extensive documentation
* open source project
* long term support
* scalable and \glref{GPU} ready
* allows fast experiments
* Python \glref{API}
\enditems

The three first points should assure that this work will be maintainable
in future. This work is faced with vast amount of data so it is desirable
to choose scalable and \glref{GPU} ready framework.
On contrary, one goal is to find great deep neural network architecture which
means that fast experimentation is crucial.
Last two points support the goal of this thesis to find the best deep neural
network architecture
and because the author of this work is comfortable with Python.

NVIDIA's website\urlnote{https://developer.nvidia.com/deep-learning-frameworks}
lists popular deep learning frameworks. These include Caffe, TensorFlow,
Theano, Torch and Keras. Torch is excluded because it has not Python
bindings \cite[torch].

\sec Caffe

Caffe is BSD-licensed C++ deep learning framework with Python and MATLAB
bindings.
It serves for training and deploying convolutional neural networks and
other deep models.
By separating model representation from actual implementation, Caffe allows
seamless switching among \glref{CPU}, \glref{GPU} and other architectures.
The project is maintained by Berkeley Vision and Learning Center with help of
community of contributors on GitHub\urlnote{https://github.com/BVLC/caffe/}.
Documentation and examples are available at Caffe
website\urlnote{http://caffe.berkeleyvision.org/}. \cite[caffe]

According to experimental results in \cite[benchmark] Caffe has one of the
best performances on both \glref{GPU} and \glref{CPU} platforms.

\sec TensorFlow

TensorFlow is library developed by Google for expressing machine learning
algorithms and an implementation for executing such algorithms.
It is licensed under Apache 2.0 open source license and its code is available on
GitHub\urlnote{https://github.com/tensorflow/tensorflow}.
A computation in TensorFlow is described by a directed graph which can be
executed with little or no change on \glref{CPU}s or \glref{GPU}s.
Currently implemented front-ends are C++ and Python.
This project has extensive documentation with examples and
tutorials\urlnote{https://www.tensorflow.org/}. \cite[tensorflow]

In order to help development and debugging TensorFlow provides visualization
tool called TensorBoard.
This tool can visualize computation graphs and summary statistics such as
values of loss function, weights and so on.

\sec Theano

Theano is a Python framework for working with mathematical expression.
Even though Theano is not developed as deep learning framework it can be
consider as it is because primarily usage is to implement
mathematical models in a symbolic way and then used them in machine or
deep learning.
Theano is open source software and is licensed under BSD license.
Its source code is versioned in repository on
GitHub\urlnote{https://github.com/Theano/Theano/}.
Main features are in language to represent mathematical expressions, a compiler to create code to compute these expressions and library to evaluate them.
This structure allows to optimize code for both \glref{CPU}s and \glref{GPU}s.
\cite[theano]

As stated in \cite[theano] although Theano is developed and mainly used for
research in deep learning it is not a deep learning framework in itself.
Several software packages have been developed to provide higher-level
user interface for instance Keras, Blocks and Lasagne.

\sec Keras

Keras is high-level neural networks \glref{API}.
It is written in Python and it runs on top of either TensorFlow or Theano.
It focuses on fast prototyping and experimentation.
Because it is interface to TensorFlow or Theano it can use both \glref{CPU}
and \glref{GPU}s.
Its website\urlnote{https://keras.io/} provides good documentation
and its code is open source under MIT license on
GitHub\urlnote{https://github.com/fchollet/keras}. \cite[keras]

Recently Keras \glref{API} is integrated into {\tt tf.contrib.keras}
module of TensorFlow 1.1.0
and then will be moved into {\tt tf.keras} in version 1.2.0.
Therefore, there will be two separate implementations of the Keras
specification one written in pure TensorFlow and second compatible with
both Theano and TensorFlow. \cite[keras2]

\sec Consideration

TensorFlow in version 1.1.0 is chosen for implementation in this thesis.
It is the most suitable framework because it has both low level interface
and high level Keras interface.
Moreover, TensorBoard is great tool for graph and dataset interactive
visualizations as it contains built-in \glref{PCA}
and \glref{t-SNE} dimensionality reduction algorithms.
Stated in \cite[karpathy2017] TensorFlow is currently the most popular
and future promising deep learning framework in machine learning community.
